Pursuing a doctoral degree during a global pandemic has been a challenging yet rewarding journey. This dissertation would not have been possible without the invaluable support and guidance of many incredible people.

First and foremost, I am deeply grateful to my advisor, Dr. Kamen Kozarev, for his insightful feedback, and the unwavering encouragement and support throughout this process. Your patience and enthusiasm have been have been truly inspirational.

My heartfelt gratitude goes to my colleagues at the Institute of Astronomy and National Astronomical Observatory, Bulgarian Academy of Sciences, especially Dr. Rositsa Miteva, Dr. Oleg Stepanyuk, Dr. Momchil Dechev, for their valuable advice, constructive feedback, and collaborative spirit.

My sincere thanks to Dr. Pietro Zucca and Dr. Peijin Zhang for introducing me to the fascinating world of radio observations from LOFAR and guiding me through the intricacies of working with the data and relevant software. Your mentorship has been invaluable.

Special thanks to my colleagues Nestor Arsenov and Yordan Darakchiev for their assistance with deep learning techniques.

I am grateful for the financial support provided by the MOSAIICS and STELLAR projects, which has allowed me to focus wholeheartedly on my research.

Words cannot express my deepest appreciation for my family, whose unwavering love, belief, and constant encouragement have been the driving force behind this achievement. To my dear friends, thank you for the laughter, beautiful memories, and companionship that have made this journey all the more fulfilling.

Finally, I acknowledge and express my heartfelt gratitude to all those who have directly or indirectly contributed to this work. Your collective efforts have played an important role in the successful completion of this dissertation.

Thank you all from the bottom of my heart.

This work was funded and supported by three projects: An ESA Contract 4000126128 to the Institute of Astronomy and NAO, BAS, under the Plan for European Cooperating States (PECS) program in Bulgaria; the Modeling and ObServational Integrated Investigations of Coronal Solar Eruptions (MOSAIICS) project, funded under contract KP-06-DV-8/18.12.2019 to the Institute of Astronomy and National Astronomical Observatory (NAO), the Bulgarian Academy of Sciences (BAS), under the National Scientific Program \textit{VIHREN} in Bulgaria; and the STELLAR project, funded by the European Union’s Horizon 2020 research and innovation program under grant agreement No. 952439.